The Effect of Events on Australian Parliamentary Discussion (1851--2017)

### 2018-10-04 outputs/docs/events/paper.Rmd
Update paper

In this paper we examine text records of what was said in Australian state and federal parliaments. Our earliest record is from 1851 and we consider the time period through to 2017. The dataset is detailed in Section \@ref(data). We use a topic model for dimensionality reduction and to allow for correlation between topics and steady changes over time. We then analyse the topics using a Bayesian hierarchical Dirichlet model to examine changes at various types of events. These events include: changes in government or elections; changes in economic conditions; and other significant events such as the 9/11 attacks. The models are detailed in Section \@ref(model).

Section \@ref(results) details our results, but broadly we find: 1) elections are associated with topic changes only when the party in power changes; 2) a change in government can be associated with topic changes even if the party in power does not change; 3) economic events, such as financial crises, have less significant and persistent effects than other events such as terrorist attacks. **[More here on the nature of the differences.]**

In this paper we apply a topic model to a dataset of larger-scale parliamentary text records from multiple Australian parliaments and then use a Bayesian hierarchical Dirichlet model to examine events. From a data perspective we contribute **[X]**. From a methods perspective, we contribute **[Y]**. Our model ican be used in many settings where the results of topic models need to be further analysed.

Our work is cross-disciplinary and in the traditions of political economy, economic history, political science and statistics. It contributes a growing modern quantitative social sciences literature that considers text as an input to more usual techniques, rather than requiring separate analysis. As the digitisation of historical sources continues and computational power becomes cheaper, we expect interest in this approach to only increase.

@@

## Analysis model
We are interested in the effect of political, economic, and other events on Australian parliamentary discussion. Political events are those related to a change of government or an election. Economic events are defined by substantial changes in various economic measures, such as the onset of the Great Depression or floating the currency. Other events are those such as the 9/11 attacks. The full list of events that we consider are detailed in Appendix \@ref(eventdetails).

@@

One output of topic modelling is a probability distribution over the topics for each document. In our case the documents are each day's text. We analyse events by focusing on those proportions, $\gamma$, which we index by sitting period $s$, time $t$ and topic $p$, that is $\gamma_{stp}$. Note that time $t$ refers to the day index within a sitting period which is any collection of days separated from other days by at least a week. Our analysis model is:

\begin{align}
\gamma_{stp} &\sim \mbox{Dirichlet}(\mu_{stp}) (\#eq:distributionforgammas) \\ 
\log \mu_{stp} &= \alpha_{sp} + \delta_{stp}. (\#eq:meanforgammas) \\
\alpha_{s1} &= 0 \\
\alpha_{sp} &\sim \mbox{Normal}\left(\eta_{g[s]}, \sigma^2_g\right) \mbox{ for } p>1 \\
\delta_{s,t,p} &\sim \mbox{Normal}\left(\delta_{s,t-1,p}, \sigma^2_{\delta}\right) \\
\sigma^2_{\delta} &\sim \mbox{Uniform}(0, 40) 
\end{align}

Equation \@ref(eq:distributionforgammas) describes the proportion of each day within a sitting period given to a particular topic as a draw from a Dirichlet distribution parameterised by $\mu_{stp}$. The Dirichlet distribution is the multivariate extension of the Beta distribution, and the parameter controls where the 'weight' of the distribution lies. Equation \@ref(eq:meanforgammas) describes that distribution parameter, $\mu_{stp}$, as the sum of $\alpha_{sp}$ which is **X**, and $\delta_{stp}$, which is **Y**. These are themselves parameters. The primary aim of the modelling task is to find an appropriate and meaningful functional form for $\mu$. We consider $\alpha_{sp}$ as a draw from a Normal distribution with a mean that depends on **A** and a variance that depends on **B**. Similarly, $\delta_{stp}$ is modelled as a draw from a Normal distribution with a mean that depends on the preceding $\delta$ and some variance which is drawn from the Uniform distribution.

@@

An illustration of the validity of our analysis model using simulated data is in Appendix \@ref(modeldetailsandsimulation).

@@

>>>>

In this paper we examine text records of what was said in Australian state and federal parliaments. Our earliest record is from 1851 and we consider the time period through to 2017. 
<!-- The dataset is detailed in Section \@ref(data).  -->
We use a topic model for dimensionality reduction and to allow for correlation between topics and steady changes over time. We then analyse the topics using a Bayesian hierarchical Dirichlet model to examine changes at various types of events. These events include: changes in government or elections; changes in economic conditions; and other significant events such as the 9/11 attacks. 
<!-- The models are detailed in Section \@ref(model). -->

<!-- Section \@ref(results) details our results, but broadly  -->
We find: 1) elections are associated with topic changes only when the party in power changes; 2) a change in government can be associated with topic changes even if the party in power does not change; 3) economic events, such as financial crises, have less significant and persistent effects than other events such as terrorist attacks. **[More here on the nature of the differences.]**

In this paper we apply a topic model to a dataset of larger-scale parliamentary text records from multiple Australian parliaments and then use a Bayesian hierarchical Dirichlet model to examine events. From a data perspective we contribute **[?????]**. From a methods perspective, we contribute **[?????]**. Our model can be used in many settings where the results of topic models need to be further analysed.

Our work is cross-disciplinary and in the traditions of political economy, economic history, political science and statistics. It contributes to a growing modern quantitative social sciences literature that considers text as an input to more traditional methods, rather than requiring separate analysis. As the digitisation of historical sources continues and computational power becomes cheaper, we expect interest in this approach to only increase.

@@

## Analysis model

@@

The main output of the STM that is of interest to us is the share of each topic in each document. We would like to know how those shares change given events. But with around 80 **[UPDATE]** topics for each of the roughly 54,331 **[UPDATE]** sitting days across the parliaments and houses the data are still too noisy to easily visualise changes around events. Instead, we put together a Bayesian hierarchical Dirichlet model that takes those shares as an input.

We index each proportion, $\gamma$, by sitting period $s$, time $t$ and topic $p$, that is $\gamma_{stp}$. Note that time $t$ refers to the day index within a sitting period which is any collection of days separated from other days by at least a week. Our analysis model is:

\begin{align}
\gamma_{stp} &\sim \mbox{Dirichlet}(\mu_{stp}) (\#eq:distributionforgammas) \\ 
\log \mu_{stp} &= \alpha_{sp} + \delta_{stp} (\#eq:meanforgammas) \\
\alpha_{s1} &= 0 (\#eq:zeroconstraint)\\
\alpha_{sp} &\sim \mbox{Normal}\left(\eta_{g[s]}, \sigma^2_g\right) \mbox{ for } p>1 (\#eq:alphasp) \\
\delta_{s,t,p} &\sim \mbox{Normal}\left(\delta_{s,t-1,p}, \sigma^2_{\delta}\right) (\#eq:deltasp) \\
\sigma^2_{\delta} &\sim \mbox{Uniform}(0, 40). (\#eq:randomwalkvariance)  
\end{align}

@@

Equation \@ref(eq:distributionforgammas) describes the proportion of each day within a sitting period given to a particular topic as a draw from a Dirichlet distribution parameterised by $\mu_{stp}$. The Dirichlet distribution is the multivariate extension of the Beta distribution, and the parameter controls where the 'weight' of the distribution lies. Equation \@ref(eq:meanforgammas) describes that distribution parameter, $\mu_{stp}$, as the sum of $\alpha_{sp}$, which means the model is linear in time for each topic and sitting period, and $\delta_{stp}$, which is a simple time series that allows for a little variation within a sitting period. These are themselves parameters. The primary aim of the modelling task is to find an appropriate and meaningful functional form for $\mu$.

Equation \@ref(eq:zeroconstraint) constrains the first intercept to zero for identification purposes. Equation \@ref(eq:alphasp) shows how we consider $\alpha_{sp}$ as a draw from a Normal distribution with a mean that depends on **[SOMETHING]** and a variance that depends on **[SOMETHING ELSE]**. Similarly, Equation \@ref(eq:deltasp) show how we model $\delta_{stp}$  as a draw from a Normal distribution with a mean that depends on the preceding $\delta$ and some variance which is drawn from the Uniform distribution in Equation \@ref(eq:randomwalkvariance), that is, a random walk around each intercept.

We run the model in Stan using the \texttt{RStan} package of @RStan. An illustration of the validity of our analysis model using simulated data is in Appendix \@ref(modeldetailsandsimulation).

@@

Political events are those related to a change of government or an election. Economic events are defined by substantial changes in various economic measures, such as the onset of the Great Depression or floating the currency. Other events are those such as the 9/11 attacks. The full list of events that we consider are detailed in Appendix \@ref(eventdetails).
